---
title: 'Prevalence of depression, before and after the COVID-19 pandemic'
output:
  officedown::rdocx_document:
    keep_md: no
csl:          ../../www/apa-old-doi-prefix.csl
bibliography: ../../www/Subestudio_confinamiento.bib
editor_options: 
  chunk_output_type: console
params:
  extra: no
---

```{r setup, include=FALSE}
library(pacman)
p_load(knitr)


# File system constants:

OUT_DIR  <- "output/COVID-19_lockdown"
SRC_DIR  <- "src"
SRC_FILE <- file.path(SRC_DIR, "Lockdown_analysis.R")


# Knitr output configuration:

opts_chunk$set(
  echo       = FALSE,
  results    = 'asis',
  warning    = FALSE,
  message    = FALSE,
  cache      = FALSE,
  fig.width  = 16.59/2.54,
  fig.height = 4.5
)

options(knitr.kable.NA = '')
options(digits = 3)
```


```{r read-chunks, cache=FALSE}
read_chunk(SRC_FILE)
```


```{r script-configuration, cache=FALSE}
```


```{r includes, cache=FALSE}
```


```{r constants}
```


```{r load-data}
```


# Prevalence of depression (Before and after the beginning of the pandemic)

```{r compute-interview-dates}
```


```{r subset-cases}
```


```{r compare-excluded-cases}
```


## Measurement of depression

```{r preprocess-data}
```


```{r create-var-labels}
```


Depression was computed with an algorithm accounting for several factors
and symptoms,
according to the Composite International Diagnostic Interview (CIDI)
for Depression Screening [@kessler_world_2004].
The Pre-measure algorithm of depression was adapted to be equivalent and
used only the same abbreviated item set that was used in the Post measure.
These items addressed depressive symptoms in the last 12 months or in the
last 30 days por the Pre and Post measures, respectively.


## Data analysis

```{r set-flextable-wd, cache=FALSE}
# This is necessary for flextable to work:
opts_knit$set(root.dir = file.path(getwd(), OUT_DIR))
```


```{r missing-responses}
```


```{r time-varying-predictors}
```


```{r subset-predictors}
```


```{r standardize-predictors}
```


```{r subset-predictors-new}
```


```{r var-descriptives}
```


Population prevalence estimates were computed
for depression both in the Pre and the Post measure.
These estimates were based on the values of the outcome variable
and the sampling weights.
The prevalence estimates were also disaggregated by sex (Male / Female)
and age group (18-49, 50+).
The Pre-measure prevalence was estimated with the sample of participants
available, before filtering out the ones excluded from the Post measure.


# Results

## Sample descriptives

- Initial sample size: `r n_initial`.

- Exclusions from the Pre measure: `r n_proxy_pre` proxy respondents.

- Sample size in the Pre measure: `r n_pre_measure`.

- Exclusions from the Post measure:

  - Participants without recontact information: `r n_missing`.
  
  - Unreachable participants: `r n_unreachable`
  (wrong number, fax line, non-answering, out of order line,
  calls repeatedly answered by an answering machine).
  
  - Deceased: `r n_deceased`.
  
  - Proxy respondents: `r n_proxy_post`.
  
  - Rejected to respond / aborted interview before finishing: `r n_reject`.
  
  - Unspecified incidents: `r n_incident`.
  

- Sample size in the post measure: `r n_final`.


## Prevalence rates

```{r depression-dataset}
```


```{r depression-descriptives}
```


```{r depression-prevalence-estimates}
```

<br>

```{r depression-prevalence-pre-table, tab.cap = "Prevalence estimates, Pre measure"}
depression_prevalence_pre %>% flextable() %>% autofit()
```

<br>

```{r depression-prevalence-post-table, tab.cap = "Prevalence estimates, Post measure"}
depression_prevalence_post %>% flextable() %>% autofit()
```

\newpage


```{r session-info, results='hide'}
```

# References
