---
title: 'Impact of the COVID-19 pandemic in mental health in Spain'
output:
  officedown::rdocx_document:
    keep_md: no
    reference_docx: ../www/Epidemiology_and_Psychiatric_Sciences.docx
csl:          ../www/who-europe-harvard.csl
bibliography: ../www/Subestudio_confinamiento.bib
editor_options: 
  chunk_output_type: console
params:
  extra: no
---

```{r setup, include=FALSE}
library(pacman)
p_load(knitr)


# File system constants:

ROOT_DIR <- ".."
OUT_DIR  <- "output"
SRC_DIR  <- "src"
SRC_FILE <- file.path(SRC_DIR, "Lockdown_analysis.R")


# opts_knit$set(root.dir = ROOT_DIR)

# Knitr output configuration:

opts_chunk$set(
  echo       = FALSE,
  results    = 'asis',
  warning    = FALSE,
  message    = FALSE,
  cache      = TRUE,
  fig.width  = 16.59/2.54,
  fig.height = 4.5
)

options(digits = 3)
```


```{r read-chunks, cache=FALSE}
read_chunk(SRC_FILE)
```


```{r script-configuration, cache=FALSE}
```


```{r includes, cache=FALSE}
```


```{r constants}
```


```{r load-data}
```


# Method

## Sample and recruitment

Non-institutionalized adults (i.e., 18+ years old)
from the regions of Madrid and Barcelona participated in this study.
These constitute the refreshment sample of the
`r PROJECT_NAME` project
(ageingandhealth.com) [@miret_health_2014].
They were recruited following a multistage stratified design consisting of:
(i) a random sample of municipalities,
with sampling probability proportional to population size;
(ii) a random sample of census units from each municipality;
(iii) a random sample of households within each census track,
delimited using predefined routes with a jump factor,
which were assigned to one of two mutually exclusive age groups:
18-49, or 50+, being the second one oversampled;
(iv) a roster of occupants was created for each household;
individuals belonging to the assigned age group were invited to participate.
The response rate was 68.0%.
Sampling weights were generated for the sample to be representative
of the target population,
according to the population distribution obtained from
the National Institute of Statistics.

```{r compute-interview-dates}
```


```{r subset-cases}
```


```{r compare-excluded-cases}
```


Participants were interviewed at their homes between
`r date_init_pre` and `r date_end_pre` (Pre-COVID measure).
They were reached out again between
`r date_init_post` and `r date_end_post`
to respond to a telephone interview (Post measure).
Trained interviewers conducted the Pre- and Post-measure interviews,
using a Computer-Assisted Personal and Telephonic Interviewing system,
respectively.
Protocols were approved by the Clinical Research Ethics Review Committees of
both Parc Sanitari Sant Joan de Déu (Barcelona),
and Hospital Universitario La Princesa (Madrid).
All participants provided informed consent for their participation and
the treatment of their personal data.

Some participants were unable to respond first-hand to one or both of
the interviews due to some physical and/or mental limitation,
and thus a relative or co-habitant answered in their name.
Only first-hand respondents to both interviews were included in these analyses;
therefore, out of a sample of `r n_initial` participants,
`r n_proxy_pre` proxy respondents were discarded,
making a sample of `r n_pre_measure` participants in the Pre measure.
A total of `r n_tota_excluded` were excluded from the Post measure
for the following reasons:
`r n_missing` participants did not provide recontact information,
`r n_unreachable` participants could not be contacted,
`r n_deceased` were deceased,
`r n_proxy_post` were responded by a proxy respondent,
`r n_reject` either rejected to respond to the Post-measure telephone interview
or aborted it before finishing,
and `r n_incident` had unspecified incidents.
The final sample comprised `r n_final` participants in the Post measure.


## Measures

```{r preprocess-data}
```


```{r create-var-labels}
```


The outcome variables investigated were depression and suicidal ideation.
Depressive symptoms were assessed with an adapted version of the
Composite International Diagnostic Interview (CIDI)
for Depression Screening [@kessler_world_2004].
An algorithm following the ICD-10 criteria
was used to diagnose depression in the previous 12 months,
although no criteria were applied to differentially diagnose
other plausible disorders.
For the Post-measure interview,
an abbreviated version was used,
and the items were adapted to ask for a 30-day time span
in order to account for an onset while
the lockdown measures were in effect.
The assessment algorithm in the Pre measure was adapted to be equivalent to
that in the Post measure,
with the same abbreviated item set.
Suicidal ideation was measured as a single item of the CIDI instrument
asking whether the participant had had suicidal thoughts
in the previous 12 months and last 30 days for the Pre- and Post-measures
periods, respectively.

The following covariates were also measured:
age, sex,
education level (coded as `r edu_level_cats`),
whether the participant lived alone (both before and during the lockdown),
whether the participant had co-habited/was co-habiting with a relative
isolated by COVID-19 (`r labels_abbr["rel_isolated"]`),
whether the participant had been/was concerned about a relative/friend
infected by COVID-19 (`r labels_abbr["rel_concerned"]`),
whether the participant had been infected with COVID-19 and its severity
(`r labels_abbr["severity"]`, coded as `r severity_cats`),
whether the participant had enough quietness at home to get proper rest
(`r labels_abbr["quietness"]`),
whether the household economic situation had worsened due to the COVID-19
emergency (`r labels_abbr["economy"]`),
whether the participant had been (temporarily) unemployed due to the COVID-19
emergency (`r labels_abbr["unemployment"]`),
time a day (in hours) spent in front of screens during the lockdown
(working and non-working),
Pre- and Post-measure levels of physical activity
coded as `r physical_cats`,
according to an abbreviated version of the
Global Physical Activity Questionnaire version 2
(GPAQ-2) [@armstrong_development_2006],
and the following scales:
Post-measure score in the Brief Resilience scale
[@rodriguez-rey_reliability_2016] measured as a 0-100 score,
Pre and Post measures of social support, measured as a 0-100 score
in the OSLO3 Social Support scale [@dalgard_negative_2006],
Pre and Post measures of loneliness measured as a 0-100 score
in the UCLA loneliness scale [@hughes_short_2016],
and Post measure of disability assessed as a 0-100 score in
the 12-item World Health Organization Disability Assessment Schedule
(WHODAS 2.0) [@luciano_12-item_2010].
These scales have been validated in previous studies (as referenced);
the Brief Resilience Scale was taken from the validated version by
Rodriguez Rey et al. [-@rodriguez-rey_reliability_2016],
while the rest of them are validated in the original in English
and were adapted for their use in the `r PROJECT_NAME` cohort study.
All of them had internal consistency indices (i.e. Cronbach's $\alpha$)
above .70, except for the OSLO3 Social Support scale,
which reliability was moderate
($\alpha$ = `r alpha_social_support_pre`)
and low ($\alpha$ = `r alpha_social_support_post`)
in the Pre and Post measures, respectively.
Despite this, we regarded these scores as important covariates in
relationship with mental health status,
and thus we decided to take them into account nevertheless.


## Data analysis

```{r set-flextable-wd, cache=FALSE}
# This is necessary for flextable to work:
opts_knit$set(root.dir = file.path(getwd(), OUT_DIR))
```


```{r missing-responses}
```


```{r time-varying-predictors}
```


```{r standardize-predictors}
```


```{r subset-predictors}
```


```{r var-descriptives}
```


Sample descriptive statistics were computed for
depression, suicidal ideation, and all the covariates.
The effect of attrition in the Pre-measure sample was analyzed
for differences in sociodemographics and the two outcome variables:
Sex, depression, and suicidal ideation were tested with the
$\mathrm{\chi}^2$-test; bias-corrected Cramér's V ($\mathrm{\phi_c}$)
was computed as a measure of effect size.
For age, a two-sample T-test was performed,
with Hedges' g as a measure of effect size.

Prevalence estimates --population-wise and disaggregated by sex
and age (grouped in `r AGE_GROUPS_OUTPUT` year-olds) --
were computed for depression and suicidal ideation in both measures.
These estimates were based on the complete data of each outcome variable
for the corresponding measure and their sampling weights.
The differences between both measures were tested with
a weighted McNemar's test of symmetry, using the complete data in both measures.
Bonferroni correction was applied variable-wise to the disaggregated estimates.

To model the risk of incidence after the lockdown,
the cases with depression or suicidal ideation in the Pre-measures period
were filtered out from the dataset for its corresponding analysis.
Then,
we performed a weighted logistic regression model on the Post measure.
All covariates stated in section Measures were initially considered.
In the case of suicidal ideation, the Pre and Post measures of depression were
also considered as covariates.
The following procedure was applied for fitting the models:
First, all interval-level variables were standardized,
in order to achieve better numerical convergence.
Categorical covariates that yielded complete separation
[@albert_existence_1984] were discarded for the same reason.
Covariates were tested individually by fitting
a univariate weighted logistic regression model
and applying the Rao-Scott Likelihood-ratio test [@rao_chi-squared_1984].
The significant covariates were selected for the
subsequent multivariate regression model
(without correction, in order to decrease Type-II error risk).
Among the selected covariates,
the ordinal ones were tested for non-linearity
comparing the general model with a model with the linear term only,
using the Wald test.
Whenever the test was non-significant, only the linear term was included.
Afterwards, a multivariate weighted logistic model was fit with the
selected covariates, and a backward-step procedure was run in order to 
drop covariates using the AIC statistic as criterion.
Covariates were dropped until the best fitting model in the next step
did not have a smaller AIC than the current step.
Demographic variables Sex and Age were considered relevant covariates and
thus were not subjected to this dropping.
In the model of suicidal ideation, the Pre and Post measures of depression
were also used as fixed covariates.
The cases with non-missing values in all the initially selected covariates
were used,
but the final model was then refit to the subset of complete cases
in the covariates selected by the backward-step procedure.
As the procedure may select
a different subset of covariates for each model,
the number of complete cases may also differ.

A significance level of `r sig_level_print` was used throughout.
All significance tests were performed
applying Bonferroni correction for multiple comparisons
(unless stated otherwise).
All the analyses were conducted in R v. 4.0.3 [@r_core_team_r:_2019].
Package survey v. 4.0 [@lumley_analysis_2004]
was used to fit the models.


# Results

## Sample descriptives

Participants with data in both measures differed from the ones
excluded in the Post measure in sex
(`r attr_sex_test_format`) and age (`r attr_age_test_format`):
The proportion of men excluded (`r prop_males_excluded`) was relatively higher
than the ones included (`r prop_males_included`).
Furthermore, the participants excluded were older
(mean = `r mean_age_excluded`, sd = `r sd_age_excluded`)
than the ones included
(mean = `r mean_age_included`, sd = `r sd_age_included`).
However, the effect size was negligible in the first case
($\mathrm{\phi_c}$ = `r attr_sex_cramerV`),
and `r attr_age_hedges_g_magnitude` in the second
(g = `r attr_age_hedges_g_format`).
No significant differences were found in depression
between the included and the excluded sample
(`r attr_depr_test_format`, $\mathrm{\phi_c}$ = `r attr_depr_cramerV`).
Regarding suicidal ideation,
the excluded sample differed significantly from the included one
(`r prop_suic_excluded` versus `r prop_suic_included`, respectively;
`r attr_suic_test_format`);
the effect size was also negligible though
($\mathrm{\phi_c}$ = `r attr_suic_cramerV`).
The descriptive statistics for both outcome variables and the covariates
for the sample included in the analysis are shown in Table 1.


## Prevalence rates

```{r depression-dataset}
```


```{r depression-prevalence-estimates}
```


```{r depression-association-population}
```


```{r suicidal-dataset}
```


```{r suicidal-prevalence-estimates}
```


```{r suicidal-association-population}
```


```{r prevalence-estimates-collapsed}
```


```{r prevalence-values}
```


Estimated prevalence rates are given in Table 2.
The estimated prevalence rate of depression increased from
`r depression_prev_pre` in the Pre to `r depression_prev_post`
in the Post measure.
According to the McNemar's test, the difference was not significant
(`r mcnemar_depr_test`).
When considering the differentiated age groups,
the difference was more prominent for the `r AGE_GROUPS[1]`
(increasing from `r depression_prev_pre_young` to
`r depression_prev_post_young`)
and the `r AGE_GROUPS[2]`
(increasing from `r depression_prev_pre_middle` to
`r depression_prev_post_middle`), but none of them was significant.
However, a significant decrease from `r depression_prev_pre_old` to
`r depression_prev_post_old` was observed in the `r AGE_GROUPS[3]` group
(`r mcnemar_depr_test_old`).
For suicidal ideation, the prevalence rate estimate increased from
`r suicidal_prev_pre` in the Pre to `r suicidal_prev_post` in the Post measure,
but this difference was not significant (`r mcnemar_suic_test`).
After Bonferroni correction,
none of the disaggregated estimates was significant either.


## Risk of depression after the lockdown

```{r depression-selection-no-pre}
```


```{r depression-no-pre-contingency}
```


```{r univariate-tests-depression}
```


```{r predictor-selection-depression}
```


```{r ordinal-linearity-tests-depression}
```


```{r stepwise-glm-depression, results='hide'}
```


```{r conclusions-pre-computations-depression}
```


```{r depression-coefficients}
```


The final regression model for depression was fit
with a sample size of `r nobs`.
Its covariates are given in Table 3,
along with their coefficients in odds-ratios.
After Bonferroni correction,
the coefficient for Age was found to be significant,
along with the Post-measures of `r depression_sig_terms_labels`.


The odds-ratio for Age was `r depr_age_OR` (`r depr_age_test`).
That is, for each year,
the risk of developing depression was expected to `r depr_age_dir`
by a `r depr_age_decr`.
For `r depr_ucla_lon_post_lab`, the odds-ratio was `r depr_ucla_lon_post_OR`
(`r depr_ucla_lon_post_test`),
which means that an increase of one standard deviation
in the Post measure of the UCLA Loneliness Scale was associated with
`r depr_ucla_lon_post_dir` of `r depr_ucla_lon_post_incr`
in the odds-ratio of receiving a positive diagnosis
of depression in the Post measure.
This corresponds to `r depr_ucla_lon_post_dir` of
`r depr_ucla_lon_post_coef_abs_score` per point in the original scoring of the
UCLA Loneliness Scale.
In the case of `r depr_resilience_post_lab`,
the odds-ratio was `r depr_resilience_post_OR` (`r depr_resilience_post_test`),
meaning that an increase of one standard deviation in the Post measure of the
Brief Resilience Scale was associated with
`r depr_resilience_post_dir` of `r depr_resilience_post_decr`
in the odds-ratio of receiving a positive diagnosis
of depression in the Post measure.
This is equivalent to `r depr_resilience_post_dir` of a
`r depr_resilience_post_coef_abs_score` per point in the original scoring of the
Brief Resilience Scale.


## Risk of suicidal ideation after the lockdown

```{r suicidal-selection-no-pre}
```


```{r suicidal-no-pre-contingency}
```


```{r univariate-tests-suicidal}
```


```{r predictor-selection-suicidal}
```


```{r drop-complete-separation-terms-suicidal}
```


```{r ordinal-linearity-tests-suicidal}
```


```{r stepwise-glm-suicidal, results='hide'}
```


```{r conclusions-pre-computations-suicidal}
```


```{r suicidal-coefficients}
```


This model was fit with a sample size of `r nobs`;
its covariates and their coefficients (as odds-ratios)
are shown in Table 4.
After Bonferroni correction,
the significant covariates were the Post-measures of
`r suicidal_sig_terms_labels_collapsed`.
The odds-ratio of `r suic_oslo3_post_label` was
`r suic_oslo3_post_coef` (`r suic_oslo3_post_test`).
This implies that an increase of one standard deviation
in the Post measure of the Oslo-3 Social Support Scale was associated with
`r suic_oslo3_post_dir` of `r suic_oslo3_post_abs`
in the odds-ratio of reporting having suicidal ideation in the Post measure or,
equivalently, `r suic_oslo3_post_dir` of a
`r suic_oslo3_post_neg_OR` per point in the original scoring of the
scale.
The odds-ratio for `r suic_resilience_post_label`
was `r suic_resilience_post_coef` (`r suic_resilience_post_test`),
implying `r suic_resilience_post_dir` of `r suic_resilience_post_abs`
for an increase of one standard deviation in the Post measure of the
Brief Resilience Scale.
This is equivalent to `r suic_resilience_post_dir` of practically
`r suic_resilience_post_neg_OR` per point in the original scoring of the
Brief Resilience Scale.

```{r session-info, results='hide'}
```

# References
