---
title: 'Impact of the COVID-19 pandemic in mental health in Spain'
output:
  officedown::rdocx_document:
    keep_md: no
    reference_docx: ../www/Epidemiology_and_Psychiatric_Sciences.docx
csl:          ../www/who-europe-harvard.csl
bibliography: ../www/Subestudio_confinamiento.bib
editor_options: 
  chunk_output_type: console
params:
  extra: no
---

```{r setup, include=FALSE}
library(pacman)
p_load(knitr)


# File system constants:

ROOT_DIR <- ".."
OUT_DIR  <- "output"
SRC_DIR  <- "src"
SRC_FILE <- file.path(SRC_DIR, "Lockdown_analysis.R")


# opts_knit$set(root.dir = ROOT_DIR)

# Knitr output configuration:

opts_chunk$set(
  echo       = FALSE,
  results    = 'asis',
  warning    = FALSE,
  message    = FALSE,
  cache      = TRUE,
  fig.width  = 16.59/2.54,
  fig.height = 4.5
)

options(digits = 3)
```


```{r read-chunks, cache=FALSE}
read_chunk(SRC_FILE)
```


```{r script-configuration, cache=FALSE}
```


```{r includes, cache=FALSE}
```


```{r constants}
```


```{r load-data}
```


# Method

## Sample and recruitment

Non-institutionalized adults (i.e., 18+ years old)
from the regions of Madrid and Barcelona participated in this study.
These constitute the refreshment sample of the
`r PROJECT_NAME` project
(ageingandhealth.com) [@miret_health_2014].
They were recruited following a multistage stratified design consisting of:
(i) a random sample of municipalities
(sampling probability proportional to population size);
(ii) a random sample of census units from each municipality;
(iii) a random sample of households within each census track,
and assigned to one of two age groups:
18-49, or 50+ (the second one oversampled).
For each household,
individuals in the assigned age group were invited to participate;
the response rate was 68.0%.
Sampling weights were generated for the sample to be representative
of the target population,
according to the population distribution obtained from
the National Institute of Statistics.

```{r compute-interview-dates}
```


```{r subset-cases}
```


```{r compare-excluded-cases}
```


Participants were interviewed at their homes between
`r date_init_pre` and `r date_end_pre` (Pre-COVID measure).
They were reached out again between
`r date_init_post` and `r date_end_post`
to respond to a telephone interview (Post measure).
Trained interviewers conducted the Pre- and Post-measure interviews,
using a Computer-Assisted Personal and Telephonic Interviewing system,
respectively.
Protocols were approved by the Clinical Research Ethics Review Committees of
Parc Sanitari Sant Joan de Déu (Barcelona)
and Hospital Universitario La Princesa (Madrid).
All participants provided informed consent.

Some participants were unable to respond first-hand
due to physical and/or mental limitations,
and thus a relative or co-habitant answered in their name.
Only first-hand respondents to both interviews were included in these analyses;
therefore, out of a sample of `r n_initial` participants,
`r n_proxy_pre` proxy respondents were discarded,
making a sample of `r n_pre_measure` participants in the Pre measure.
A total of `r n_tota_excluded` were excluded from the Post measure
(`r n_missing` participants did not provide recontact information,
`r n_unreachable` participants could not be contacted,
`r n_deceased` were deceased,
`r n_proxy_post` were responded by a proxy respondent,
`r n_reject` either rejected to respond to the Post-measure telephone interview
or aborted it before finishing,
and `r n_incident` had unspecified incidents),
so the final Post-measure sample comprised `r n_final` participants.


## Measures

```{r preprocess-data}
```


```{r create-var-labels}
```


Depressive symptoms were assessed with an adapted version of the
Composite International Diagnostic Interview (CIDI)
for Depression Screening [@kessler_world_2004].
An algorithm following the ICD-10 criteria
was used to diagnose depression in the previous 12 months
[@world_health_organization_icd-10_1992].
For the Post-measure interview,
an abbreviated version was used,
and the items were adapted to ask for a 30-day time span
in order to account for an onset while
the lockdown measures were in effect.
The assessment algorithm in the Pre measure was adapted to use the same item set
as in the Post measure.
Suicidal ideation comprised a single item
asking whether the participant had had suicidal thoughts
in the previous 12 months / 30 days, for the Pre- and Post-measures
periods, respectively.

The following covariates were also measured:
age, sex,
education level,
whether the participant lived alone(both before and during the lockdown),
whether the participant had co-habited/was co-habiting with a relative
isolated by COVID-19,
whether the participant had been/was concerned about a relative/friend
infected by COVID-19,
whether the participant had been infected with COVID-19 and its severity,
whether the participant had enough quietness at home to get proper rest,
whether the household economic situation had worsened due to the COVID-19,
whether the participant had been unemployed due to the COVID-19,
time a day spent in front of screens during the lockdown
(working and non-working),
Pre- and Post-measure levels of physical activity
according to an abbreviated version of the
Global Physical Activity Questionnaire version 2
(GPAQ-2) [@armstrong_development_2006],
and the following scales:
Post-measure score in the Brief Resilience scale
[@rodriguez-rey_reliability_2016],
Pre and Post measures of social support measured with
in the OSLO3 Social Support scale [@dalgard_negative_2006],
Pre and Post measures of loneliness measured with
the UCLA loneliness scale [@hughes_short_2016],
and Post measure of disability assessed with
the 12-item World Health Organization Disability Assessment Schedule
(WHODAS 2.0) [@luciano_12-item_2010].
The Brief Resilience Scale was taken from the validated version by
Rodriguez Rey et al. [@rodriguez-rey_reliability_2016],
while the rest of them have been validated in the original in English
(as referenced)
and were adapted for their use in the `r PROJECT_NAME` cohort study.
All of them had internal consistency indices (i.e., Cronbach's $\alpha$)
above .70, except for the OSLO3 Social Support scale,
which reliability was moderate
($\alpha$ = `r alpha_social_support_pre`)
and low ($\alpha$ = `r alpha_social_support_post`)
in the Pre and Post measures, respectively.


## Data analysis

```{r set-flextable-wd, cache=FALSE}
# This is necessary for flextable to work:
opts_knit$set(root.dir = file.path(getwd(), OUT_DIR))
```


```{r missing-responses}
```


```{r time-varying-predictors}
```


```{r standardize-predictors}
```


```{r subset-predictors}
```


```{r var-descriptives}
```


Sample descriptive statistics were computed for
depression, suicidal ideation, and all the covariates.
Attrition in the Pre-measure sample was analyzed
for differences in sociodemographics and the two outcome variables:
Sex, depression, and suicidal ideation were tested with the
$\mathrm{\chi}^2$-test; bias-corrected Cramér's V ($\mathrm{\phi_c}$)
was computed as a measure of effect size.
For age, a two-sample T-test was performed,
with Hedges' g as a measure of effect size.

Prevalence estimates --population-wise and disaggregated by sex
and age (grouped in `r AGE_GROUPS_OUTPUT` year-olds) --
were computed for depression and suicidal ideation in both measures.
The differences between both measures were tested with
a weighted McNemar's test of symmetry, using the complete data.
Bonferroni correction was applied variable-wise to the disaggregated estimates.

To model the risk of incidence after the lockdown,
the cases with depression or suicidal ideation in the Pre-measures period
were filtered out from the dataset for its corresponding analysis.
Then,
we performed a weighted logistic regression model on the Post measure.
All covariates stated in section Measures were initially considered.
In the case of suicidal ideation, the Pre and Post measures of depression were
also considered as covariates.
The following procedures were applied for fitting the models:
First,
in order to archive better numerical convergence,
all interval-level variables were standardized,
and categorical covariates that yielded complete separation
[@albert_existence_1984] were discarded.
Covariates were tested individually with
univariate weighted logistic regression models
and the Rao-Scott [-@rao_chi-squared_1984] Likelihood-ratio test
(without Bonferroni correction, in order to decrease Type-II error risk).
Among the significant covariates,
the ordinal ones were tested for non-linearity with the Wald test,
comparing the general model with a model with the linear term only.
Whenever the test was non-significant, only the linear term was included.
Afterwards, a multivariate weighted logistic model was fit with all the
significant covariates. A backward-step procedure was then run,
dropping covariates according to the AIC statistic.
Demographic variables Sex and Age were fixed, excluding them from dropping.
In the model of suicidal ideation, the measures of depression
were also fixed.
Finally, the resulting model was refit to the subset of complete cases
in the covariates selected by the backward-step procedure.
As the procedure may select
a different subset of covariates for each model,
the number of complete cases may also differ.

A significance level of `r sig_level_print` was used throughout.
All significance tests were performed
applying Bonferroni correction for multiple comparisons
(unless stated otherwise).
All the analyses were conducted in R v. 4.0.3 [@r_core_team_r:_2019].
Package survey v. 4.0 [@lumley_analysis_2004]
was used to fit the models.


# Results

## Sample descriptives

Participants with data in both measures differed from the ones
excluded in the Post measure in sex
(`r attr_sex_test_format`) and age (`r attr_age_test_format`):
The proportion of men excluded (`r prop_males_excluded`) was relatively higher
than the ones included (`r prop_males_included`),
and the participants excluded were older
(mean = `r mean_age_excluded`, sd = `r sd_age_excluded`)
than the ones included
(mean = `r mean_age_included`, sd = `r sd_age_included`).
However, the effect size was negligible for sex
($\mathrm{\phi_c}$ = `r attr_sex_cramerV`),
and `r attr_age_hedges_g_magnitude` for age
(g = `r attr_age_hedges_g_format`).
No significant differences were found in depression
between the included and the excluded sample
(`r attr_depr_test_format`, $\mathrm{\phi_c}$ = `r attr_depr_cramerV`).
Regarding suicidal ideation,
the excluded sample differed significantly from the included one
(`r prop_suic_excluded` versus `r prop_suic_included`, respectively;
`r attr_suic_test_format`);
the effect size was also negligible though
($\mathrm{\phi_c}$ = `r attr_suic_cramerV`).
The descriptive statistics for both outcome variables and the covariates
for the sample included in the analysis are shown in Table 1.


## Prevalence rates

```{r depression-dataset}
```


```{r depression-prevalence-estimates}
```


```{r depression-association-population}
```


```{r suicidal-dataset}
```


```{r suicidal-prevalence-estimates}
```


```{r suicidal-association-population}
```


```{r prevalence-estimates-collapsed}
```


```{r prevalence-values}
```


Estimated prevalence rates are given in Table 2.
For depression, it increased from
`r depression_prev_pre` in the Pre to `r depression_prev_post`
in the Post measure.
According to the McNemar's test, the difference was not significant
(`r mcnemar_depr_test`).
When considering the differentiated age groups,
the difference was more prominent for the `r AGE_GROUPS[1]`
(increasing from `r depression_prev_pre_young` to
`r depression_prev_post_young`)
and the `r AGE_GROUPS[2]`
(increasing from `r depression_prev_pre_middle` to
`r depression_prev_post_middle`) groups, but none of them was significant.
However, a significant decrease from `r depression_prev_pre_old` to
`r depression_prev_post_old` was observed in the `r AGE_GROUPS[3]` group
(`r mcnemar_depr_test_old`).
For suicidal ideation, the prevalence rate estimate increased from
`r suicidal_prev_pre` in the Pre to `r suicidal_prev_post` in the Post measure,
but this difference was not significant (`r mcnemar_suic_test`).
After Bonferroni correction,
none of the disaggregated estimates was significant either.


## Risk of depression after the lockdown

```{r depression-selection-no-pre}
```


```{r depression-no-pre-contingency}
```


```{r univariate-tests-depression}
```


```{r predictor-selection-depression}
```


```{r ordinal-linearity-tests-depression}
```


```{r stepwise-glm-depression, results='hide'}
```


```{r conclusions-pre-computations-depression}
```


```{r depression-coefficients}
```


The final regression model for depression was fit
with a sample size of `r nobs`.
Its covariates are given in Table 3,
along with their odds-ratios.
The coefficient for Age was found to be significant,
along with the Post-measures of `r depression_sig_terms_labels`.


The odds-ratio for Age was `r depr_age_OR` (`r depr_age_test`);
for each year,
the risk of developing depression was expected to `r depr_age_dir`
by `r depr_age_decr`.
For `r depr_ucla_lon_post_lab`, the odds-ratio was `r depr_ucla_lon_post_OR`
(`r depr_ucla_lon_post_test`),
which means that an increase of one standard deviation
in the Post measure of the UCLA Loneliness Scale was associated with
`r depr_ucla_lon_post_dir` of `r depr_ucla_lon_post_incr`
in the odds-ratio of receiving a positive diagnosis
of depression in the Post measure.
In the case of `r depr_resilience_post_lab`,
the odds-ratio was `r depr_resilience_post_OR` (`r depr_resilience_post_test`),
meaning that an increase of one standard deviation in the Post measure of the
Brief Resilience Scale was associated with
`r depr_resilience_post_dir` of `r depr_resilience_post_decr`
in the odds-ratio of receiving a positive diagnosis
of depression in the Post measure.


## Risk of suicidal ideation after the lockdown

```{r suicidal-selection-no-pre}
```


```{r suicidal-no-pre-contingency}
```


```{r univariate-tests-suicidal}
```


```{r predictor-selection-suicidal}
```


```{r drop-complete-separation-terms-suicidal}
```


```{r ordinal-linearity-tests-suicidal}
```


```{r stepwise-glm-suicidal, results='hide'}
```


```{r conclusions-pre-computations-suicidal}
```


```{r suicidal-coefficients}
```


This model was fit with a sample size of `r nobs`;
its covariates and their coefficients (as odds-ratios)
are shown in Table 4.
After Bonferroni correction,
the significant covariates were the Post-measures of
`r suicidal_sig_terms_labels_collapsed`.
The odds-ratio of `r suic_oslo3_post_label` was
`r suic_oslo3_post_coef` (`r suic_oslo3_post_test`).
This implies that an increase of one standard deviation
in the Post measure of the Oslo-3 Social Support Scale was associated with
`r suic_oslo3_post_dir` of `r suic_oslo3_post_abs`
in the odds-ratio of reporting having suicidal ideation in the Post measure.
The odds-ratio for `r suic_resilience_post_label`
was `r suic_resilience_post_coef` (`r suic_resilience_post_test`),
implying `r suic_resilience_post_dir` of `r suic_resilience_post_abs`
for an increase of one standard deviation in the Post measure of the
Brief Resilience Scale.

```{r session-info, results='hide'}
```

# References
