---
title: 'Baseline loneliness associated with post depression and/or suicidal ideation'
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: kable
  officedown::rdocx_document:
    reference_docx: ../www/Epidemiology_and_Psychiatric_Sciences.docx
csl:          ../www/who-europe-harvard.csl
bibliography: ../www/Subestudio_confinamiento.bib
editor_options: 
  chunk_output_type: console
params:
  extra: no
---

```{r setup, include=FALSE}
library(pacman)
p_load(knitr)


# File system constants:

ROOT_DIR <- ".."
OUT_DIR  <- "output"
SRC_DIR  <- "src"
SRC_FILE <- file.path(SRC_DIR, "Lockdown_analysis.R")


opts_knit$set(root.dir = ROOT_DIR)

# Knitr output configuration:

opts_chunk$set(
  echo       = FALSE,
  results    = 'asis',
  warning    = FALSE,
  message    = FALSE,
  cache      = TRUE,
  fig.width  = 16.59/2.54,
  fig.height = 4.5
)

options(digits = 3)
```


# Intro

First, we source all the code from the paper, to make sure the results are
the same as published.
This gives tables 1 through 4 as output.
(This tables are fromatted as the Word output, so are not html optimized.)

```{r read-chunks, cache=FALSE}
read_chunk(SRC_FILE)
```


```{r script-configuration, cache=FALSE}
```


```{r includes, cache=FALSE}
```


```{r constants}
```


```{r load-data}
```


```{r compute-interview-dates}
```


```{r subset-cases}
```


```{r compare-excluded-cases}
```


```{r preprocess-data}
```


```{r create-var-labels}
```


```{r set-flextable-wd, cache=FALSE}
# This is necessary for flextable to work:
opts_knit$set(root.dir = file.path(getwd(), OUT_DIR))
```


```{r missing-responses}
```


```{r time-varying-predictors}
```


```{r standardize-predictors}
```


```{r subset-predictors}
```


```{r var-descriptives}
```


```{r depression-dataset}
```


```{r depression-prevalence-estimates}
```


```{r depression-association-population}
```


```{r suicidal-dataset}
```


```{r suicidal-prevalence-estimates}
```


```{r suicidal-association-population}
```


```{r prevalence-estimates-collapsed}
```


```{r prevalence-values}
```


```{r depression-selection-no-pre}
```


```{r depression-no-pre-contingency}
```


```{r univariate-tests-depression}
```


```{r predictor-selection-depression}
```


```{r ordinal-linearity-tests-depression}
```


```{r stepwise-glm-depression, results='hide'}
```


```{r conclusions-pre-computations-depression}
```


```{r depression-coefficients}
```


```{r suicidal-selection-no-pre}
```


```{r suicidal-no-pre-contingency}
```


```{r univariate-tests-suicidal}
```


```{r predictor-selection-suicidal}
```


```{r drop-complete-separation-terms-suicidal}
```


```{r ordinal-linearity-tests-suicidal}
```


```{r stepwise-glm-suicidal, results='hide'}
```


```{r conclusions-pre-computations-suicidal}
```


```{r suicidal-coefficients}
```


# Original tables

## Table 1

```{r descriptive-stats-table, tab.id="descriptive-stats-table", tab.cap=CAPTION_DESC_VARIABLES, cache=FALSE}
total_descriptives_out
```


## Table 2

```{r prevalence-table, tab.id="prevalence-table", tab.cap=CAPTION_PREVALENCES, cache=FALSE}
prevalence_table_output
```


## Table 3

```{r depression-coefficients-table, tab.id="depression-coefficients-table", tab.cap=CAPTION_DEPRESSION_FIT_NEW_TERMS, cache=FALSE}
depression_coefficients_table
```


## Table 4

```{r suicidal-coefficients-table, tab.id="suicidal-coefficients-table", tab.cap=CAPTION_SUICIDAL_FIT_NEW_TERMS, cache=FALSE}
suicidal_coefficients_table
```


# Information requested

## OR for baseline loneliness associated with post depression

We compute this by adding the pre-loneliness score to
the logistic regression model for post-lockdown depression.

```{r, echo=TRUE}
depression_fit_pre_loneliness <- depression_fit |>
  update(". ~ . + ucla_lon_pre")
```


```{r, results='markup'}
depression_fit_pre_loneliness |>
  tidy() |>
  mutate(
    OR         = exp(estimate) %>% number(1e-2),
    ci.inf     = exp(estimate - std.error * CI_FACTOR),
    ci.sup     = exp(estimate + std.error * CI_FACTOR),
    `(95% CI)` = format_ci(ci.inf, ci.sup, sig = 2, quoting = "("),
    `*p* value`  = p.value   %>% format_pvalues(),
    sig        = p.value %>% is_less_than(SIG_LEVEL / (n() - 1)) %>%
      if_else("*", ""), # -1 to omit intercept
    statistic = statistic %>% number(1e-2),
    across(where(is.numeric), number, 1e-3)
  ) %>%
  format_term_label(
    .data   = depression_neg_pre,
    .labels = var_properties %>% pull(labels_comp),
    add_ref = FALSE
  ) %>%
  order_terms_with_data(dataset_outcomes) |>
  select(Term, OR, `(95% CI)`, `*p* value`)
```


## OR for baseline loneliness associated with suicidal ideation

The same is done for the model for post-lockdown suicidal ideation.

```{r, echo=TRUE}
suicidal_fit_pre_loneliness <- suicidal_fit |>
  update(". ~ . + ucla_lon_pre")
```


```{r, results='markup'}
suicidal_fit_pre_loneliness |>
  tidy() |>
  mutate(
    OR         = exp(estimate) %>% number(1e-2),
    ci.inf     = exp(estimate - std.error * CI_FACTOR),
    ci.sup     = exp(estimate + std.error * CI_FACTOR),
    `(95% CI)` = format_ci(ci.inf, ci.sup, sig = 2, quoting = "("),
    `*p* value`  = p.value   %>% format_pvalues(),
    sig        = p.value %>% is_less_than(SIG_LEVEL / (n() - 1)) %>%
      if_else("*", ""), # -1 to omit intercept
    statistic = statistic %>% number(1e-2),
    across(where(is.numeric), number, 1e-3)
  ) %>%
  format_term_label(
    .data   = depression_neg_pre,
    .labels = var_properties %>% pull(labels_comp),
    add_ref = FALSE
  ) %>%
  order_terms_with_data(dataset_outcomes) |>
  select(Term, OR, `(95% CI)`, `*p* value`)
```

## Univariate regression models

In order to provide further information, we estimate the univariate
models and tabulate their coefficient.

### Depression

```{r, echo=TRUE}
depression_fit_pre_loneliness_only <- depression_fit |>
  update(". ~ ucla_lon_pre")
```


```{r, results='markup'}
depression_fit_pre_loneliness_only |>
  tidy() |>
  mutate(
    OR         = exp(estimate) %>% number(1e-2),
    ci.inf     = exp(estimate - std.error * CI_FACTOR),
    ci.sup     = exp(estimate + std.error * CI_FACTOR),
    `(95% CI)` = format_ci(ci.inf, ci.sup, sig = 2, quoting = "("),
    `*p* value`  = p.value   %>% format_pvalues(),
    sig        = p.value %>% is_less_than(SIG_LEVEL / (n() - 1)) %>%
      if_else("*", ""), # -1 to omit intercept
    statistic = statistic %>% number(1e-2),
    across(where(is.numeric), number, 1e-3)
  ) %>%
  format_term_label(
    .data   = depression_neg_pre,
    .labels = var_properties %>% pull(labels_comp),
    add_ref = FALSE
  ) %>%
  order_terms_with_data(dataset_outcomes) |>
  select(Term, OR, `(95% CI)`, `*p* value`)
```


### Suicidal ideation

```{r, echo=TRUE}
suicidal_fit_pre_loneliness_only <- suicidal_fit |>
  update(". ~ ucla_lon_pre")
```


```{r, results='markup'}
suicidal_fit_pre_loneliness_only |>
  tidy() |>
  mutate(
    OR         = exp(estimate) %>% number(1e-2),
    ci.inf     = exp(estimate - std.error * CI_FACTOR),
    ci.sup     = exp(estimate + std.error * CI_FACTOR),
    `(95% CI)` = format_ci(ci.inf, ci.sup, sig = 2, quoting = "("),
    `*p* value`  = p.value   %>% format_pvalues(),
    sig        = p.value %>% is_less_than(SIG_LEVEL / (n() - 1)) %>%
      if_else("*", ""), # -1 to omit intercept
    statistic = statistic %>% number(1e-2),
    across(where(is.numeric), number, 1e-3)
  ) %>%
  format_term_label(
    .data   = depression_neg_pre,
    .labels = var_properties %>% pull(labels_comp),
    add_ref = FALSE
  ) %>%
  order_terms_with_data(dataset_outcomes) |>
  select(Term, OR, `(95% CI)`, `*p* value`)
```


# Sessio info

```{r session-info, results='markup'}
```
