---
title: "COVID-19 substudy database management"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(pacman)
p_load(haven, tidyverse, knitr)
```


# Intro

This notebook describes the process of conforming and merging the databases
from the COVID-19 substudy within the Edad con Salud study.


# Input data

There are two master databases, one with the data collected from
participants in Madrid and the other one in Catalonia.
These data files are encrypted with a password; `haven` package's functions to
read SPSS data files do not implement the password feature, so the files are
manually opened in SPSS, and then saved into the project `dat` folder,
with the names
`dataset_surveys_COVID_followup_MAD.sav` and
`dataset_surveys_COVID_followup_CAT.sav`.

```{r}
DATA_PATH <- file.path("dat")

DB_MAD_PATH <- file.path(DATA_PATH, "dataset_surveys_COVID_followup_MAD.sav")
DB_CAT_PATH <- file.path(DATA_PATH, "dataset_surveys_COVID_followup_CAT.sav")

db_mad <- DB_MAD_PATH %>% read_sav()
db_cat <- DB_CAT_PATH %>% read_sav()
```


After reading these datasets, they are immediately deleted:

```{r, results='hide'}
file.remove(DB_MAD_PATH, DB_CAT_PATH)
```


# Data processing

We only need to merge the two datasets.
However, not all the variable names are equal:

```{r}
tibble(MAD = colnames(db_mad), BCN = colnames(db_cat)) %>% filter(MAD != BCN)
```

This variable name has to be equated accross both databases before collapsing
them:

```{r}
db_mad <- db_mad %>% rename(ESTADO_ENTREVISTA = ESTADO_ENTREV)
```

Then we collapse de datasets, adding variable `origin` to identify the dataset
where each variable comes originally from (MAD / BCN):

```{r}
db_list <- list(
  MAD = db_mad,
  BCN = db_cat
)

db_out <- bind_rows(db_list, .id = "origin")

db_out <- db_out %>% mutate(
  origin = origin %>%
    factor()      %>%
    labelled(
      labels = set_names(sort(unique(as.integer(.))), levels(.)),
      label  = "Origin dataset"
    )
)
```


Then we need to delete the sensitive variables (i.e. with personal information):

```{r}
SENSITIVE_VARS <- c(
  "TELEFONO", "TEL_UTILIZADO", "NOMBRES", "CPOSTAL_BD", "MUNICIPIO", "CONTACTO",
  "TLF2", "NOMBRE_PROXY"
)

db_out <- db_out %>% select(-all_of(SENSITIVE_VARS))
```


# Write data

The complete dataset is written both in SPSS and Stata format.

```{r}
OUTPUT_PATH <- file.path(
  "~/../UAM",
  "marta.miret@uam.es - Bases de datos maestras Edad con Salud",
  "Subestudio_COVID"
)

OUTPUT_FILE <- file.path(OUTPUT_PATH, "Edad_con_salud_Fichero_Completo")

db_out %>% write_sav(OUTPUT_FILE %>% paste0(".sav"))
db_out %>% write_dta(OUTPUT_FILE %>% paste0(".dta"))
```
