---
title: "Physical activity in Edad con Salud (COVID lockdown substudy)"
output:
  html_notebook:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# Packages:
library(pacman)
p_load(tidyverse, haven, psych, magrittr, broom)
```


```{r constants, include=FALSE}
# File system:
BASE_DIR <- "~/../UAM"
DB_PATH_MASTER <- file.path(
  BASE_DIR,
  "marta.miret@uam.es - Bases de datos maestras Edad con Salud"
)
DOC_PATH_MASTER <- file.path(
  BASE_DIR,
  "marta.miret@uam.es - Documentacion Edad con Salud"
)
OUTCOME_PATH_BASE <- file.path(
  DOC_PATH_MASTER,
  "Edad con salud - Ola 3/Outcomes/Cohorte 2019",
  "Submuestra_1_preconfinamiento/Outcome datasets"
)
OUTCOME_PATH_COVID <- file.path(
  DOC_PATH_MASTER,
  "Edad con salud - Subestudio COVID/Outcomes/Outcome datasets"
)
OUTPUT_PATH <- file.path(
  DOC_PATH_MASTER,
  "Edad con salud - Subestudio COVID/BBDD_con_outcomes"
)

# Collapsed outcomes file:
OUTPUT_FILE <- file.path(OUTPUT_PATH, "Outcomes_collapsed.dta")

# Variables:
BASELINE_ID_VARS <- c("number_id", "q0002_hhid")
POST_LD_ID <- "ID_CONTACTO"
POST_LD_ID_VARS  <- c(POST_LD_ID, BASELINE_ID_VARS)
```


We are finding a very high rate of participants with "High" levels of
physical activity, according to the
[WHO's GPAQ criteria][1].
We want to explore the Pre and Post dataset, in order to identify where these
differences come from.

[1]: http://www.who.int/chp/steps/GPAQ%20%20%20%20%20%20%20%20Instrument%20and%20Analysis%20Guide%20v2.pdf

# Loading the datasets

```{r load-master-datasets}
dataset_master_pre  <- file.path(
  DB_PATH_MASTER,
  "Ola_3/Cohorte_2019/Submuestra_1_preconfinamiento",
  "Edad con Salud ola 3_cohorte 2019_base completa_Stata14.dta"
) %>%
  read_dta()

dataset_master_post <- file.path(
  DB_PATH_MASTER,
  "Subestudio_COVID",
  "Edad_con_salud_Fichero_Completo.dta"
) %>%
  read_dta() %>%
  rename(number_id = IDENTIFICA1, q0002_hhid = IDENTIFICA2)
```


# Outcome variable generation

This code is exactly equivalent to the Stata code implemented in the
Baseline wave for the 2019 cohort of the Edad con Salud study,
as appears in the outcome description documents for
[the original variable][2],
and [the version comparable with the Post measure][3].

[2]: https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Ola%203/Outcomes/Cohorte%202019/Submuestra_1_preconfinamiento/Outcome%20descriptions/Description_physical%20activity.docx?d=wa5ce4ec33ffe4455a7f08e3fa4947194&csf=1&web=1&e=ncgkOy

[3]: https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Subestudio%20COVID/Outcomes/Outcome%20descriptions/Physical%20activity_SubestudioCOVID_comparable2019.docx?d=wb855216be12843a391b989eb18484407&csf=1&web=1&e=wYiypW

```{r recode-physical-activity-pre}
dataset_physical_pre <- dataset_master_pre %>% transmute(
  
  across(all_of(BASELINE_ID_VARS)), q0007a_result, proxy,
  
  across(
    q3017_days:q3030_mins & ends_with(c("days", "hours", "mins")),
    ~if_else(is.na(.) & q0007a_result == 1, 0, .) # Missing to 0
  ),
  
  # Total time for each activity domain:
  q3018_1  = q3018_hours * 60 + q3018_mins,
  q3021_1  = q3021_hours * 60 + q3021_mins,
  q3024_1  = q3024_hours * 60 + q3024_mins,
  q3027_1  = q3027_hours * 60 + q3027_mins,
  q3030_1  = q3030_hours * 60 + q3030_mins,
  

  ## Original definition according to the GPAQ:
  
  # Total MET reckoning:
  MET1     = q3017_days * q3018_1 * 8 + # Vigorous, job
    q3020_days * q3021_1 * 4 +          # Moderate, job
    q3023_days * q3024_1 * 4 +          # Commuting
    q3026_days * q3027_1 * 8 +          # Vigorous, leisure
    q3029_days * q3030_1 * 4,           # Moderate, leisure
  
  MET1     = if_else( # Correction of unplausible MET values:
    ( #   higher than 16 hours a week in a single domain...
      q3018_1 > 960 | q3021_1 > 960 | q3024_1 > 960 | q3027_1 > 960 |
        q3030_1 > 960
    ) | #   ... OR...
      ( #   ... non-0 MET values where the number of days are 0 (coding error).
        (q3017_days == 0 & q3018_1 > 0) | (q3020_days == 0 & q3021_1 > 0) |
          (q3023_days == 0 & q3024_1 > 0) | (q3026_days == 0 & q3027_1 > 0) |
          (q3029_days == 0 & q3030_1 > 0)
      ),
    NA_real_, # Substitute these cases by missing values
    MET1      # Leave as is otherwise (if previous condition is not met)
  ),
  
  # Recoding as activity level categories: High, Moderate, Low:
  physical = case_when(
    # Exceptions:
    #   When MET1 is a "missing"
    is.na(MET1) ~ NA_character_,
    
    #   When all activities are "missing"
    is.na(q3016_vigorous) & is.na(q3019_moderate) & is.na(q3022_walk) &
      is.na(q3025_leisure) & is.na(q3028_moderate) ~ NA_character_,
    
    # Day tally of vigorous activity of 3 or higher AND 1500 METs or higher...
    ((q3017_days + q3026_days) >= 3 & (MET1 >= 1500)) |
      (#   ... or total day tally of 7 or higher AND 3000 METs or higher.
        (
          (q3017_days + q3020_days + q3023_days + q3026_days + q3029_days) >= 7
        ) &
          (MET1 >= 3000)
      ) ~ "High", # These two conditions meet the criteria for "High" activity.
    
    (
      ( # Day tally of vigorous activity of 3 or higher AND 60 METs or higher...
        (q3017_days + q3026_days) >= 3 & #   ... in vigorous activities,...
          ((q3017_days * q3018_1) + (q3026_days * q3027_1)) >= 60
      ) | #   ... OR...
        ( #   ...day tally of moderate activity of 5 or higher AND...
          (q3020_days + q3023_days + q3029_days) >= 5 &
            ( #   ...150 METs or higher in moderate activities...
              (q3020_days * q3021_1) +
                (q3023_days * q3024_1) +
                (q3029_days * q3030_1)
            ) >= 150
        ) | #   ... OR...
        ( #   ... total tally day of 5 or higher AND 600 METs or higher, ...
          (
            q3017_days + q3020_days + q3023_days + q3026_days + q3029_days
          ) >= 5 &
            (MET1 >= 600)
        )
    ) ~ "Moderate", # These conditions meet the criteria for "Moderate" activity
    
    # Any other case: "Low" activity (as default if is still "NA")
    TRUE ~ "Low"
  ) %>%
    factor(levels = c("High", "Moderate", "Low")),
  
  
  ## Comparable pre-post lockdown:
  
  # Total MET reckoning:
  MET2     = q3017_days * q3018_1 * 8 + # Vigorous, job
    q3020_days * q3021_1 * 4 +          # Moderate, job
    q3023_days * q3024_1 * 4,           # Commuting

  MET2     = if_else( # Correction of unplausible MET values:
    ( #   higher than 16 hours a week in a single domain...
      q3018_1 > 960 | q3021_1 > 960 | q3024_1 > 960
    ) | #   ... OR...
      ( #   ... non-0 MET values where the number of days are 0 (coding error).
        (q3017_days == 0 & q3018_1 > 0) | (q3020_days == 0 & q3021_1 > 0) |
          (q3023_days == 0 & q3024_1 > 0)
      ),
    NA_real_, # Substitute these cases by missing values
    MET2      # Leave as is otherwise (if previous condition is not met)
  ),
  
  physical_comp = case_when(
    # Exceptions:
    #   When MET2 is a "missing"
    is.na(MET2) ~ NA_character_,
    
    #   When all activities are "missing"
    is.na(q3016_vigorous) & is.na(q3019_moderate) & is.na(q3022_walk) ~
      NA_character_,
    
    # Day tally of vigorous activity of 3 or higher AND 1500 METs or higher...
    (q3017_days >= 3 & (MET2 >= 1500)) |
      (#   ... or total day tally of 7 or higher AND 3000 METs or higher.
        ((q3017_days + q3020_days + q3023_days) >= 7) & (MET2 >= 3000)
      ) ~ "High", # These two conditions meet the criteria for "High" activity.
    
    (
      ( # Day tally of vigorous activity of 3 or higher AND 60 METs or higher...
        q3017_days >= 3 & #   ... in vigorous activities,...
          (q3017_days * q3018_1) >= 60
      ) | #   ... OR...
        ( #   ...day tally of moderate activity of 5 or higher AND...
          (q3020_days + q3023_days) >= 5 &
            ( #   ...150 METs or higher in moderate activities...
              (q3020_days * q3021_1) + (q3023_days * q3024_1)
            ) >= 150
        ) | #   ... OR...
        ( #   ... total tally day of 5 or higher AND 600 METs or higher, ...
          (q3017_days + q3020_days + q3023_days) >= 5 & (MET2 >= 600)
        )
    ) ~ "Moderate", # These conditions meet the criteria for "Moderate" activity
    
    # Any other case: "Low" activity (as default if is still "NA")
    TRUE ~ "Low"
  ) %>%
    factor(levels = c("High", "Moderate", "Low"))
)

dataset_physical_pre %>% count(physical)
dataset_physical_pre %>% count(physical_comp)
```


This code is also equivalent to the computation in Stata of the outcome
variable for the Post measure, as appears in its [description document][4].

[4]: https://dauam-my.sharepoint.com/:w:/r/personal/marta_miret_uam_es/Documents/Edad%20con%20Salud/Documentacion%20Edad%20con%20Salud/Edad%20con%20salud%20-%20Subestudio%20COVID/Outcomes/Outcome%20descriptions/Physical%20activity_SubestudioCOVID.docx?d=wab40b0af90ad467f8584581a339c576b&csf=1&web=1&e=b2gOjX

```{r recode-physical-activity-post}
dataset_physical_post <- dataset_master_post %>% transmute(
  
  across(all_of(POST_LD_ID_VARS)), ESTADO_ENTREVISTA,
  
  across(
    AF1A:AF3B_M_1 & ends_with(c("A", "_1")),
    ~if_else(
      (is.na(.) | . == 99L) & ESTADO_ENTREVISTA == 1,
      0L, as.integer(.)
    ) # Missing to 0
  ),
  
  # Total time for each activity domain:
  AF1B_1   = AF1B_H_1 * 60 + AF1B_M_1,
  AF2B_1   = AF2B_H_1 * 60 + AF2B_M_1,
  AF3B_1   = AF3B_H_1 * 60 + AF3B_M_1,
  
  # Total MET reckoning:
  MET_post = AF1A * AF1B_1 * 8 + # Vigorous, total
    AF2A * AF2B_1 * 4 +          # Moderate, total
    AF3A * AF3B_1 * 4,           # Commuting
  
  MET_post = if_else( # Correction of unplausible MET values:
    ( #   higher than 16 hours a week in a single domain...
      AF1B_1 > 960 | AF2B_1 > 960 | AF3B_1 > 960
    ) | #   ... OR...
      ( #   ... non-0 MET values where the number of days are 0 (coding error).
        (AF1A == 0 & AF1B_1 > 0) | (AF2A == 0 & AF2B_1 > 0) |
          (AF3A == 0 & AF3B_1 > 0)
      ),
    NA_real_, # Substitute these cases by missing values
    MET_post  # Leave as is otherwise (if previous condition is not met)
  ),
  
  physical_post = case_when(
    # Exceptions:
    #   When MET_post is a "missing"
    is.na(MET_post) ~ NA_character_,
    
    # There are no equivalent items in the Post-measure interview
    # #   When all activities are "missing"
    # is.na(q3016_vigorous) & is.na(q3019_moderate) & is.na(q3022_walk) ~
    #   NA_character_,
    
    # Day tally of vigorous activity of 3 or higher AND 1500 METs or higher...
    (AF1A >= 3 & (MET_post >= 1500)) |
      (#   ... or total day tally of 7 or higher AND 3000 METs or higher.
        ((AF1A + AF2A + AF3A) >= 7) & (MET_post >= 3000)
      ) ~ "High", # These two conditions meet the criteria for "High" activity.
    
    (
      ( # Day tally of vigorous activity of 3 or higher AND 60 METs or higher...
        AF1A >= 3 & #   ... in vigorous activities,...
          (AF1A * AF1B_1) >= 60
      ) | #   ... OR...
        ( #   ...day tally of moderate activity of 5 or higher AND...
          (AF2A + AF3A) >= 5 &
            ( #   ...150 METs or higher in moderate activities...
              (AF2A * AF2B_1) + (AF3A * AF3B_1)
            ) >= 150
        ) | #   ... OR...
        ( #   ... total tally day of 5 or higher AND 600 METs or higher, ...
          (AF1A + AF2A + AF3A) >= 5 & (MET_post >= 600)
        )
    ) ~ "Moderate", # These conditions meet the criteria for "Moderate" activity
    
    # Any other case: "Low" activity (as default if is still "NA")
    TRUE ~ "Low"
  ) %>%
    factor(levels = c("High", "Moderate", "Low"))
)

dataset_physical_post %>% count(physical_post)
```


# Case filtering

We create a dataset filtering the cases, to keep only the ones that are used
in the COVID-lockdown substudy.

```{r}
dataset_physical <- dataset_physical_post %>%
  full_join(dataset_physical_pre, by = c("number_id", "q0002_hhid")) %>%
  filter(ESTADO_ENTREVISTA == 1, proxy == 2)

dataset_physical %>% count(physical)
dataset_physical %>% count(physical_comp)
dataset_physical %>% count(physical_post)
```


# Exploration of the differences in the distributions

## Exploration of the distributions in days, hours and minutes of each activity domain

We compute the statistics of the original acitivity domain variables, to see
whether a difference in the distribution may explain these differences.

```{r}
dataset_physical %>%
  mutate(across(ends_with("_1"), na_if, 99L)) %>%
  select(
    q3017_days:q3030_mins & ends_with(c("days", "hours", "mins")),
    AF1A:AF3B_M_1 & ends_with(c("A", "_1"))
  ) %>%
  describe(skew = FALSE)
```


The `max` values show that there may be some coding errors in the values.
However, these do not explain the difference in the distributions.
They may occur both in the Pre and the Post interview variables.

There is a relevant increase in the number of days dedicated to moderate
intensity activities.
In the Pre measure, the mean is .83 days for moderate-intensity working 
activities (`q3017_days`) and 1.02 for moderate-intensity leisure-time
activities (`q3029_days`).
In the Post measure, the mean of total days with moderate-intensity activities
is 4.89 (`AF2A`).


## Distribution tests of days, hours and minutes

We compute the equivalent number of days, hours and minutes for vigorous and
moderate-intensity activities, by adding the values of the working and
leisure-time settings.

```{r}
dataset_physical <- dataset_physical %>%
  mutate(
    vigorous_days_pre  = q3017_days  + q3026_days,
    vigorous_hours_pre = q3018_hours + q3027_hours,
    vigorous_mins_pre  = q3018_mins  + q3027_mins,
    moderate_days_pre  = q3020_days  + q3029_days,
    moderate_hours_pre = q3021_hours + q3030_hours,
    moderate_mins_pre  = q3021_mins  + q3030_mins
  )
```


Now we may compare the distributions by means of a paired t-test for each pair
of variables (in the Pre and the Post), testing whether the Post measure
is greater than the Pre measure:

```{r}

vars <- tribble(
  ~Domain,           ~Pre,                 ~Post,
   "Days vigorous",   "vigorous_days_pre",  "AF1A",
   "Hours vigorous",  "vigorous_hours_pre", "AF1B_H_1",
   "Mins vigorous",   "vigorous_mins_pre",  "AF1B_M_1",
   "Days moderate",   "moderate_days_pre",  "AF2A",
   "Hours moderate",  "moderate_hours_pre", "AF2B_H_1",
   "Mins moderate",   "moderate_mins_pre",  "AF2B_M_1",
   "Days commuting",  "q3023_days",         "AF3A",
   "Hours commuting", "q3024_hours",        "AF3B_H_1",
   "Mins commuting",  "q3024_mins",         "AF3B_M_1",
)

vars %$% map2(
  Post, Pre,
  ~t.test(
    dataset_physical[[.x]],
    dataset_physical[[.y]],
    paired      = TRUE,
    alternative = "greater"
  ) %>%
    tidy()
) %>%
  map2(vars %>% pull(Domain), ~add_column(.x, variable = .y, .before = 1)) %>%
  bind_rows()
```


There is a very significant increase, both in days and
in time per day of moderate activity.
There also seems that there could be a decrease of time in hours and days of
vigorous activity.


## Distribution tests of total time per week

We will see now if the total time per week is different in the Pre and Post
measure.

```{r}
dataset_physical <- dataset_physical %>% mutate(
  vigorous_time_pre   = q3017_days * q3018_1 + q3026_days * q3027_1,
  moderate_time_pre   = q3020_days * q3021_1 + q3029_days * q3030_1,
  commuting_time_pre  = q3023_days * q3024_1,
  vigorous_time_post  = AF1A * AF1B_1,
  moderate_time_post  = AF2A * AF2B_1,
  commuting_time_post = AF3A * AF3B_1
)

vars2 <- tribble(
  ~Domain,              ~Pre,                 ~Post,
  "Time/week vigorous",  "vigorous_time_pre",  "vigorous_time_post",
  "Time/week moderate",  "moderate_time_pre",  "moderate_time_post",
  "Time/week commuting", "commuting_time_pre", "commuting_time_post"
)

vars2 %$% map2(
  Post, Pre,
  ~t.test(dataset_physical[[.x]], dataset_physical[[.y]], paired = TRUE) %>%
    tidy()
) %>%
  map2(vars2 %>% pull(Domain), ~add_column(.x, variable = .y, .before = 1)) %>%
  bind_rows()
```

There is a decrease of 79.4 minutes of vigorous activity per week,
and also a decrease of 64.5 minutes of activity while commuting, both being
significative.
However, it is much more than compensated by an increase of 422 minutes of
moderate activity per week, 7 hours per week.


## Distribution test of METs

The resulting total METs for each measure are:

```{r}
dataset_physical <- dataset_physical %>% mutate(
  METs_pre  = vigorous_time_pre * 8 +
    moderate_time_pre  * 4 +
    commuting_time_pre * 4,
  METs_post = vigorous_time_post * 8 +
    moderate_time_post  * 4 +
    commuting_time_post * 4
)

dataset_physical %$% t.test(METs_post, METs_pre, paired = TRUE) %>%
  tidy() %>%
  mutate(
    eq_vigorous_time = estimate / 8,
    eq_moderate_time = estimate / 4
  )
```

The resulting difference in METs is also significant, amounting to 793
(equivalent to 99 minutes of vigorous exercise,
or 198 minutes of moderate exercise).


# Explanations of the differences

The difference of the distributions in physical activity categories
(Low, Moderate, High) is clearly due to the difference in METs in the
sample.
The thresholds and criteria for considering what are Moderate- and
High-intensity physical activiy, according to the GPAQ algorithm,
result in a large amount of the sample shifting upwards due to a high increase
in moderate-intensity activity time, which in turn result in a significant
increase of total amount of METs per week.

There are several, non-excluding, possible explanations of the increase in
moderate-intensity activity time in the sample:

- **Coding errors**: It is not impossible that the variables for minutes and
hours have been mistaken in the data entry process.
The distributions of these variables should be reviewed.
The data entry process may be double-checked again, in order to look for
possible errors in the interview protocols, or discrepancies between the
protocols and the database.

- **Changes in the measurement properties**: It is possible that the response
criterion has changed in the population due to the lockdown circumstances.
This could be due to a more relaxed criteria setting the thresholds for the
self-assessment of the number of days and time per day dedicated to
each activity, or could be due to a social desirability
(i.e. self-deception or impression-management) bias in the response,
given that higher levels of physical activity are probably regarded as more
socially acceptable. There is no way of testing this hypothesis with
the present dataset and instruments.

- **A genuine increase in moderate-intensity activity time**: It is possible
that the population represented in the sample has genuinely increased their
activity time, as a result of having more spare time available for exercising.
The decrease in vigorous-activity time may imply that some participants which
exercise regularly may have had to change their routines to adapt to the
hardware and capabilities they had during the lockdown, thus increasing the
time of moderate activity, overcompensating for the loss of vigorous-activity
time. Finally, it is likely that some people may have had started
exercising as a means of avoiding boredom and/or lockdown-related anxiety.


# Update (28/10/2020): Interpretation of the items

After finding there may be an issue with the formulation of the item stems,
we hypothesize that the items AF2b and AF3b may have been understood as
"total activity time in the last 30 days" (and not activity time per day).

We compare the results by dividing the time in these two items by 30, to get
what would be a daily equivalent.

```{r}
dataset_physical_post_alt <- dataset_master_post %>% transmute(
  
  across(all_of(POST_LD_ID_VARS)), ESTADO_ENTREVISTA,
  
  across(
    AF1A:AF3B_M_1 & ends_with(c("A", "_1")),
    ~if_else(
      (is.na(.) | . == 99L) & ESTADO_ENTREVISTA == 1,
      0L, as.integer(.)
    ) # Missing to 0
  ),
  
  # Total time for each activity domain:
  AF1B_1   = AF1B_H_1 * 60 + AF1B_M_1,
  AF2B_1   = (AF2B_H_1 * 60 + AF2B_M_1) / 30,
  AF3B_1   = (AF3B_H_1 * 60 + AF3B_M_1) / 30,
  
  # Total MET reckoning:
  MET_post = AF1A * AF1B_1 * 8 + # Vigorous, total
    AF2A * AF2B_1 * 4 +          # Moderate, total
    AF3A * AF3B_1 * 4,           # Commuting
  
  MET_post = if_else( # Correction of unplausible MET values:
    ( #   higher than 16 hours a week in a single domain...
      AF1B_1 > 960 | AF2B_1 > 960 | AF3B_1 > 960
    ) | #   ... OR...
      ( #   ... non-0 MET values where the number of days are 0 (coding error).
        (AF1A == 0 & AF1B_1 > 0) | (AF2A == 0 & AF2B_1 > 0) |
          (AF3A == 0 & AF3B_1 > 0)
      ),
    NA_real_, # Substitute these cases by missing values
    MET_post  # Leave as is otherwise (if previous condition is not met)
  ),
  
  physical_post = case_when(
    # Exceptions:
    #   When MET_post is a "missing"
    is.na(MET_post) ~ NA_character_,
    
    # There are no equivalent items in the Post-measure interview
    # #   When all activities are "missing"
    # is.na(q3016_vigorous) & is.na(q3019_moderate) & is.na(q3022_walk) ~
    #   NA_character_,
    
    # Day tally of vigorous activity of 3 or higher AND 1500 METs or higher...
    (AF1A >= 3 & (MET_post >= 1500)) |
      (#   ... or total day tally of 7 or higher AND 3000 METs or higher.
        ((AF1A + AF2A + AF3A) >= 7) & (MET_post >= 3000)
      ) ~ "High", # These two conditions meet the criteria for "High" activity.
    
    (
      ( # Day tally of vigorous activity of 3 or higher AND 60 METs or higher...
        AF1A >= 3 & #   ... in vigorous activities,...
          (AF1A * AF1B_1) >= 60
      ) | #   ... OR...
        ( #   ...day tally of moderate activity of 5 or higher AND...
          (AF2A + AF3A) >= 5 &
            ( #   ...150 METs or higher in moderate activities...
              (AF2A * AF2B_1) + (AF3A * AF3B_1)
            ) >= 150
        ) | #   ... OR...
        ( #   ... total tally day of 5 or higher AND 600 METs or higher, ...
          (AF1A + AF2A + AF3A) >= 5 & (MET_post >= 600)
        )
    ) ~ "Moderate", # These conditions meet the criteria for "Moderate" activity
    
    # Any other case: "Low" activity (as default if is still "NA")
    TRUE ~ "Low"
  ) %>%
    factor(levels = c("High", "Moderate", "Low"))
)

dataset_physical_post_alt %>% count(physical_post)
```

The resulting distribution of activities reflects this fact.


## Statistics of the physical activity variables

```{r}
dataset_physical %>%
  mutate(
    across(ends_with("_1"), na_if, 99L),
    across(c("AF2B_H_1", "AF2B_M_1", "AF3B_H_1", "AF3B_M_1"), `/`,   30)
  ) %>%
  select(
    q3017_days:q3030_mins & ends_with(c("days", "hours", "mins")),
    AF1A:AF3B_M_1 & ends_with(c("A", "_1"))
  ) %>%
  describe(skew = FALSE)
```

The values are now extremely low. It is unclear though whether these
are plausible (they are expected to be lower, compared to the Pre measure)
are they are unrealistically low.


## Distribution tests of days, hours and minutes

```{r}
dataset_physical <- dataset_physical %>%
  mutate(
    moderate_hours_post  = AF2B_H_1 / 30,
    moderate_mins_post   = AF2B_M_1 / 30,
    commuting_hours_post = AF3B_H_1 / 30,
    commuting_mins_post  = AF3B_M_1 / 30
  )
```


Paired t-tests:

```{r}
vars <- tribble(
  ~Domain,           ~Pre,                 ~Post,
   "Days vigorous",   "vigorous_days_pre",  "AF1A",
   "Hours vigorous",  "vigorous_hours_pre", "AF1B_H_1",
   "Mins vigorous",   "vigorous_mins_pre",  "AF1B_M_1",
   "Days moderate",   "moderate_days_pre",  "AF2A",
   "Hours moderate",  "moderate_hours_pre", "moderate_hours_post",
   "Mins moderate",   "moderate_mins_pre",  "moderate_mins_post",
   "Days commuting",  "q3023_days",         "AF3A",
   "Hours commuting", "q3024_hours",        "commuting_hours_post",
   "Mins commuting",  "q3024_mins",         "commuting_mins_post",
)

vars %$% map2(
  Post, Pre,
  ~t.test(
    dataset_physical[[.x]],
    dataset_physical[[.y]],
    paired      = TRUE
  ) %>%
    tidy()
) %>%
  map2(vars %>% pull(Domain), ~add_column(.x, variable = .y, .before = 1)) %>%
  bind_rows()
```

Hours and minutes of moderate and commuting activities are now lower in the
Post measure, as we would expect.


## Distribution tests of total time per week

```{r}
dataset_physical <- dataset_physical %>% mutate(
  vigorous_time_pre   = q3017_days * q3018_1 + q3026_days * q3027_1,
  moderate_time_pre   = q3020_days * q3021_1 + q3029_days * q3030_1,
  commuting_time_pre  = q3023_days * q3024_1,
  vigorous_time_post  = AF1A * AF1B_1,
  moderate_time_post  = (moderate_hours_post * 60 + moderate_mins_post) * AF2A,
  commuting_time_post = (commuting_hours_post * 60 + commuting_mins_post) * AF3A
)

vars2 <- tribble(
  ~Domain,              ~Pre,                 ~Post,
  "Time/week vigorous",  "vigorous_time_pre",  "vigorous_time_post",
  "Time/week moderate",  "moderate_time_pre",  "moderate_time_post",
  "Time/week commuting", "commuting_time_pre", "commuting_time_post"
)

vars2 %$% map2(
  Post, Pre,
  ~t.test(dataset_physical[[.x]], dataset_physical[[.y]], paired = TRUE) %>%
    tidy()
) %>%
  map2(vars2 %>% pull(Domain), ~add_column(.x, variable = .y, .before = 1)) %>%
  bind_rows()
```

The decrease in activity time per week is:

* 79.4 minutes ($\approx$ 1 hour 20') for vigorous-intensity activity,

* 245 minutes ($\approx$ 4 hours) for moderate-intensity activity, and

* 310  minutes ($\approx$ 5 hours) for commuting.


## Distribution test of METs

```{r}
dataset_physical <- dataset_physical %>% mutate(
  METs_pre  = vigorous_time_pre * 8 +
    moderate_time_pre  * 4 +
    commuting_time_pre * 4,
  METs_post = vigorous_time_post * 8 +
    moderate_time_post  * 4 +
    commuting_time_post * 4
)

dataset_physical %$% t.test(METs_post, METs_pre, paired = TRUE) %>%
  tidy() %>%
  mutate(
    eq_vigorous_time = estimate / 8,
    eq_moderate_time = estimate / 4
  )
```

The resulting difference in METs is a decrease of 2857.73
(equivalent to 357 minutes of vigorous exercise,
or 714 minutes of moderate exercise).


# Session info

```{r}
devtools::session_info()
```

